%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Array jumping game}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Problem}

An array has integer elements.  The integer of a cell represents the
number of positions we can jump to the right from the cell.  Starting
from the cell at index $0$, determine whether we are able to reach the
last cell of the array.  Our answer must be either $1$ (meaning true)
or $0$ (meaning false).


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Solution}

The problem statement does not say anything about whether the integers
are negative, positive, or zero.  If an array element is a negative
integer, does that mean we jump backward?  Let's make various
assumptions to simplify the problem:

\begin{packedenumeral}
\item Each integer in the array is nonnegative.
\item We start from the zeroth index of the array.
\item Each jump must be to the right, increasing the array index.
\item If an array element is zero, we cannot jump.
\end{packedenumeral}

A simple approach to this problem is of a greedy nature.  Start from
the first element of the array and jump the maximum distance we can to
reach a different array cell.  We have $2$ cases:

\begin{packedenumeral}
\item We are not yet at the last cell of the array.

\item We are at the last cell of the array.
\end{packedenumeral}

If we are at the last cell of the array, then we are done.  Otherwise
jump the maximum distance we can to reach a different array cell.

Suppose we are at an array cell that allows us to jump a maximum
distance of $d \geq 0$.  Suppose $d = 0$.  If the cell is the first
cell of the array, then we cannot reach the end of the array.
Otherwise backtrack to the cell we came from and decrease the element
of that cell by $1$, effectively reducing the maximum jump distance
allowed by the cell.  Suppose $d > 0$.  Either we can jump the given
maximum distance or we cannot.  If we can jump the given maximum
distance, then we do so to reach a different cell.  Otherwise we
backtrack and decrement as per above.  Our discussion is summarized in
\Algorithm{alg:jump:array_jumping_game}.

\begin{algorithm}[!htbp]
\input{algorithm/jump.tex}
\caption{%%
  Greedy array jump.
}
\label{alg:jump:array_jumping_game}
\end{algorithm}
